The function--sheaf dictionary for smooth commutative group schemes over finite fields, with applications over local fields

### 2014-08-22 dictionary.tex
Commit to go home. . .

Let $G$ be a commutative group scheme over $\Fq$.
Consider the category $\mathcal{C}(G)$ of triples
$\qcs{F}\ceq (\gqcs{F},\mu_\qcs{F},\phi_\qcs{F})$ where
\begin{enumerate}
\labitem{(C.0)}{C.0} $\gqcs{F}$ is a rank one $\ell$-adic local system on $\bG\ceq G \times_\Spec{\Fq} \Spec{\bFq}$,
\labitem{(C.1)}{C.1} $\mu_\qcs{F}: \bm^* \gqcs{F} \to \gqcs{F}\boxtimes\gqcs{F}$ is an isomorphism $\ell$-adic sheaves, where $\bm : \bG \times \bG \to \bG$ is multiplication $m$ for $G$ extended to $\bG$.
\labitem{(C.2)}{C.2} $\phi_\qcs{F} : \Frob{G}^* \gqcs{F} \to \gqcs{F}$ is an isomorphism of constructible $\ell$-adic sheaves, where $\Frob{G}^* : \bG  \to \bG$ is the Frobenius automorphism,
\end{enumerate}
subject to the conditions as in \ref{QC.1} and \ref{QC.2},
with morphisms defined as in \ref{QC.3}.
This category $\mathcal{C}(G)$ is rigid monoidal, with tensor product defined as in Definition~\ref{def:QC}. %Note that $\mathcal{C}(G) = \QC(G)$ if $G$ is also smooth.

The Greenberg transform $\Gr_R(X)$ of $X$ is commutative group scheme over $\Fq$ with component group scheme
$
\pi_0(\Gr_R(X))= \pi_0(X) \times_S S_0
$
and $\Fq$-rational points
$
\Gr_R(X)(\Fq)\iso X_K(K),
$
canonically.
For every non-negative integer $n$, $\Gr^R_n(X)$ is {\it smooth} commutative group scheme over $\Fq$ with component group scheme
$\pi_0(\Gr^R_n(X))= \pi_0(X) \times_S S_0$
and $\Fq$-rational points
$
\Gr^R_N(X)(\Fq)\iso X(R_n),
$
canonically.
%
Moreover, the morphism of $\Fq$-schemes $P^R_n(X) : \Gr_R(X) \to \Gr^R_n(N)$ induces a functor
\[
P^R_n(X)^* : \QC(\Gr^R_n(X)) \to \mathcal{C}(\Gr_R(X)),
\]
as in Proposition~\ref{prop:pullback}.
This functor is full and faithful, so we may view $\QC(\Gr^R_n(X))$ as a full subcategory of $\mathcal{C}(\Gr_R(X))$.

\begin{definition}
Let $X_K$ be a connected commutative group scheme that admits a Neron model, $X$.
A {\it quasicharacter sheaf on $\Gr_R(X)$} is an object $\qcs{F}$ in $\mathcal{C}(\Gr_R(X))$ such that $\qcs{F} \iso P^R_n(X)^*\qcs{L}$ for some quasicharacter sheaf $\qcs{L}$ on $\Gr^R_n(X)$.
Quasicharacter sheaves on $\Gr_R(X)$ form a rigid monoidal category in the obvious way, denoted by $\QC(\Gr_R(X))$.
\end{definition}

@@

\begin{theorem}\label{thm:QCXK}
Let $K$ and $R$ be as above.
The trace of Frobenius provides a natural transformation between the additive functors
\[
F_{\QC}: X_K \mapsto \QCiso{\Gr_R(X)}
\qquad\text{and}\qquad
F_{\text{qc}}: X_K \mapsto \Hom_{\text{qc}}(X_K(K),\EEx)
\]
where $X$ is a Néron model for $X_K$, as functors from $\mathcal{N}_K$ to the category of abelian groups. 
Regarding this natural transformation, for every $X_K \in \mathcal{N}_K$ with Néron model $X$:
\begin{enumerate}
\item[(1)] there is a canonical short exact sequence of abelian groups 
\[
0 \to \Hh^2(\pi_0(X)_{\bFq},\EEx)^{\Frob{}} \to \QCiso{\Gr_R(X)} \to \Hom_\text{qc}(X_K(K),\EEx) \to 0;
\] 
\item[(2)] for all quasicharacter sheaves $\qcs{F}$, $\qcs{F}'$ on $\Gr_R(X)$, and for every $\rho \in \Hom(\qcs{F},\qcs{F}')$, either $\rho$ is trivial or $\rho$ is an isomorphism;
\item[(3)] for all quasicharacter sheaves $\qcs{F}$ on $\Gr_R(X)$,
\[
\Aut(\qcs{F}) \iso \Hom((\pi_0(X)_{\bFq})_{\Weil{}},\EEx),
\]
canonically.
\end{enumerate}
\end{theorem}

>>>>

Let $X_K \in \mathcal{N}_K$ and $X$ be as above.
Although $\Gr_R(X)$ is not locally of finite type, and thus not smooth, let us write $\QC(\Gr_R(X))$ for the rigid monoidal category defined as in Section~\ref{sec:category}, without insisting that the commutative group scheme $G$ is smooth.

The Greenberg transform $\Gr_R(X)$ of $X$ is commutative group scheme over $\Fq$ with component group scheme
$
\pi_0(\Gr_R(X))= \pi_0(X) \times_S S_0
$
and $\Fq$-rational points
$
\Gr_R(X)(\Fq)\iso X_K(K),
$
canonically.
For every non-negative integer $n$, $\Gr^R_n(X)$ is {\it smooth} commutative group scheme over $\Fq$ with component group scheme
$\pi_0(\Gr^R_n(X))= \pi_0(X) \times_S S_0$
and $\Fq$-rational points
$
\Gr^R_N(X)(\Fq)\iso X(R_n),
$
canonically.
%
Moreover, the morphism of $\Fq$-schemes $P^R_n(X) : \Gr_R(X) \to \Gr^R_n(N)$ induces a functor
\[
P^R_n(X)^* : \QC(\Gr^R_n(X)) \to \QC(\Gr_R(X)),
\]
as in Proposition~\ref{prop:pullback}.
This functor is full and faithful, so we may view $\QC(\Gr^R_n(X))$ as a full subcategory of $\QC(\Gr_R(X))$.

\begin{definition}
Let $X_K$ be a connected commutative group scheme that admits a Neron model, $X$.
A {\it quasicharacter sheaf for $X_K$} is an object $\qcs{F}$ in $\QC(\Gr_R(X))$ such that $\qcs{F} \iso P^R_n(X)^*\qcs{L}$ for some $\qcs{L} \in \QC(\Gr^R_n(X))$.
Quasicharacter sheaves for $X_K$ form a rigid monoidal category in the obvious way, denoted by $\QC(X_K)$.
\end{definition}

@@

\begin{theorem}\label{thm:QCXK}
Let $K$ and $R$ be as above.
The trace of Frobenius provides a natural transformation between the additive functors
\[
F_{\QC}: X_K \mapsto \QCiso{X_K}
\qquad\text{and}\qquad
F_{\text{qc}}: X_K \mapsto \Hom_{\text{qc}}(X_K(K),\EEx)
\]
where $X$ is a Néron model for $X_K$, as functors from $\mathcal{N}_K$ to the category of abelian groups. 
Regarding this natural transformation, for every $X_K \in \mathcal{N}_K$ with Néron model $X$:
\begin{enumerate}
\item[(1)] there is a canonical short exact sequence of abelian groups 
\[
0 \to \Hh^2(\pi_0(X)_{\bFq},\EEx)^{\Frob{}} \to \QCiso{X_K} \to \Hom_\text{qc}(X_K(K),\EEx) \to 0;
\] 
\item[(2)] for all quasicharacter sheaves $\qcs{F}$, $\qcs{F}'$ on $\Gr_R(X)$, and for every $\rho \in \Hom(\qcs{F},\qcs{F}')$, either $\rho$ is trivial or $\rho$ is an isomorphism;
\item[(3)] for all quasicharacter sheaves $\qcs{F}$ on $\Gr_R(X)$,
\[
\Aut(\qcs{F}) \iso \Hom((\pi_0(X)_{\bFq})_{\Weil{}},\EEx),
\]
canonically.
\end{enumerate}
\end{theorem}
